\documentclass{beamer}
\usepackage{listings}
\usepackage{minted}
\usepackage{hyperref}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{animate}
\usepackage{xparse}

\newcommand\dsuml{\sum\limits}

\title{BRACU CP Workshop}
\subtitle{Day 6}
\author{
    Ahnaf Shahriar Asif\\
    Md Mubasshir Chowdhury
}

\institute{BRAC University, Dhaka}

\date{30 April 2025}

\begin{document}

\frame{\titlepage}
% Add the Table of Contents on the second slide
\begin{frame}
  \frametitle{Table of Contents}
  \tableofcontents
\end{frame}

\section{Introduction}
\begin{frame}{What will we learn ?}
\begin{itemize}
    \item<1-> Divisibility  
    \item<2-> Primes
    \item<3-> Modular Arithmetic
    \item<4-> Primality testing
    \item<5-> Number theoretic functions
\end{itemize}
\end{frame}

\section{Divisibility}
\begin{frame}{Divisibility}
\begin{itemize}
  \item<1-> Let \(a\), \(b\), and \(c\) be integers. Then:
  \begin{enumerate}[(i)]
    \item If \(b \mid a\) and \(c \mid a\), then \(bc \mid a\).
    \item If \(b \mid a\) and \(c \mid b\), then \(c \mid a\).
    \item If \(b \mid a\) and \(c \mid b\), then \(bc \mid ab\).
    \item If \(c \mid a\) and \(c \mid b\), then \(c \mid (a + b)\) and \(c \mid (a - b)\).
  \end{enumerate}
\end{itemize}
\end{frame}

\section{Finding divisors of n}
\begin{frame}{Finding divisors of n}
\begin{itemize}
  \item<1-> We can loop through all numbers and see if they divide \(n\). 
  \item<2-> Looping till $\sqrt{n}$ is enough.
  \item<3-> We can find divisors of all numbers upto $n$ in $\mathcal{O}(n \ln(n))$ time.
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Code}
  \frametitle{Finding divisors of n}
  \begin{itemize}
    \item<1-> \begin{minted}{cpp}
vector<vector<int>> divisors(n + 1);
for (int i = 1; i <= n; i++) {
    for (int j = i; j <= n; j += i) {
        divisors[j].push_back(i);
    }
}
\end{minted}

  \end{itemize}
\end{frame}

\section{Primes}
\begin{frame}[fragile]{Primes}
  \begin{itemize}
    \item<1-> A prime $p$ is greater than 1 and only divisible by 1 and $p$.
    \item<2-> A composite number is a positive integer that is not prime.
    \item<3-> Prime numbers help us calculate a lot of things easily.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Find number of divisors of $n$}
  \begin{itemize}
    \item<1-> Given an integer $1 \leq n \leq 10^9$, find the number of divisors of $n$.
  \end{itemize}
\end{frame}
\begin{frame}[fragile]{Solution}
  \frametitle{Solution: Number of divisors}
  \begin{itemize}
    \item<1-> We can find the prime factorization of $n$.
    \item<2-> If $n = p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$, then the number of divisors of $n$ is given by:
    \[
      d(n) = (e_1 + 1)(e_2 + 1) \cdots (e_k + 1)
    \]
  \end{itemize}
\end{frame}


\begin{frame}[fragile]{Prime Factorization}
  \frametitle{Prime Factorization}
  \begin{itemize}
    \item<1-> We can find the prime factorization of $n$ in $\mathcal{O}(\sqrt{n})$ time.
    \item<2-> We can use a \textbf{sieve} to find all primes up to $10^6$ in $\mathcal{O}(n \log(\log(n)))$ time.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Code}
  \frametitle{Prime Factorization code}
  \begin{minted}{cpp}
  // vector<int> primes; (contains all primes upto sqrt(n))
  vector<pair<int,int>> ppf(int n){
    vector<pair<int,int>> factors;
    for (int i = 0; primes[i] * primes[i] <= n; i++){
      int cnt = 0;
      while (n % primes[i] == 0){
        n /= primes[i];
        cnt++;
      }
      if (cnt > 0) factors.push_back({primes[i], cnt});
    }
    if (n > 1) factors.push_back({n, 1});
    return factors;
  }
  // if n = 12, it will return:
  // {{2, 2}, {3, 1}} which means 12 = 2^2 * 3^1 
  \end{minted}
\end{frame}

\begin{frame}[fragile]{Sieve}
  \frametitle{Sieve}
  \begin{itemize}
    \item<1-> We can use a sieve to find all primes up to $10^6$ in $\mathcal{O}(n \log(\log(n)))$ time.
    \item<2-> We can use the sieve to find the smallest prime factor of each number.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Sieve Simulation}
  \frametitle{Sieve Simulation}
  
\begin{center}
\begin{tabular}{|*{10}{c|}}
\hline
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
\hline
11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\
\hline
21 & 22 & 23 & 24 & 25 & 26 & 27 & 28 & 29 & 30 \\
\hline
31 & 32 & 33 & 34 & 35 & 36 & 37 & 38 & 39 & 40 \\
\hline
\end{tabular}
\end{center}
\end{frame}

\begin{frame}[fragile]{Code}
  \frametitle{Sieve Code}
  \begin{minted}{cpp}
  bool is_composite[N];
  vector<int> primes;

  void sieve(int n){
    is_composite[1] = true;
    for(int i = 2; i <= n;i++){
      if(!is_composite[i]){
        primes.push_back(i);
        for(int j = i * i; j <= n;j+=i){
          is_composite[j] = true;
        }
      }
    }
  }
\end{minted}
\end{frame}

\begin{frame}{HS08PAUL}
  \frametitle{HS08PAUL - A conjecture of Paul Erdos}
  \framesubtitle{Problem Link: \url{https://www.spoj.com/problems/HS08PAUL/}}
  \begin{itemize}
\item<1-> compute the number of \textbf{positive primes} not larger than a given integer \( n \), which can be expressed in the form \( x^2 + y^4 \) for some integers \( x \) and \( y \).
  \item<2-> There will be $10^4$ test cases, and $1 \leq n \leq 10^6$.
  \end{itemize}
\end{frame}

\begin{frame}{Rough}
  \frametitle{Rough - HS08PAUL}
\end{frame}
\begin{frame}{Solution}
  \begin{itemize}
    \item<1-> 
First of all, we can find all the primes upto $10^7$ using sieve. Let's say we've the \textbf{is\_composite} array. We haven't pushed anything in the \textbf{primes} vector yet. Now, what we can do is go through all possible $x^2+y^4$ and mark all the primes that are of that form. It is easy to notice that $x^2 \leq 10^7$ and $y^4 \leq 10^7$. Or in other words, $x \leq 10^4$ and $y \leq 60$. So, we can definitely loop through all possible $x$ and $y$, and for each pair, we can check if $x^2+y^4$ is prime. If it is, we can mark it in the \textbf{is\_composite} array. Based on that array, we can generate a prefix sum array that can answer all the queries. And we can do all of these before the queries. 
  \end{itemize}
\end{frame}
\section{GCD, LCM}
\begin{frame}{GCD, LCM}
  \begin{itemize}
    \item<1-> GCD: Greatest Common Divisor
    \item<2-> LCM: Least Common Multiple
    \item<3-> GCD(a, b) = GCD(b, a mod b)
    \item<4-> LCM(a, b) = (a * b) / GCD(a, b)
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{GCD}
  \frametitle{GCD}
  \begin{itemize}
    \item<1-> GCD(a, b) can be computed in $\mathcal{O}(\log(\min(a, b)))$ time.
    \item<2-> Code:
        \begin{minted}{cpp}
          int gcd(int a, int b){
            if (b == 0)return a;
            return gcd(b, a % b);
          }
        \end{minted}
    \item<3-> if $a = p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$ and $b = p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$, then:
        \[
          GCD(a, b) = p_1^{\min(e_1, f_1)} \cdot p_2^{\min(e_2, f_2)} \cdots p_k^{\min(e_k, f_k)}
        \]
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{LCM}
  \frametitle{LCM}
  \begin{itemize}
    \item<1-> LCM(a, b) can be computed in $\mathcal{O}(\log(\min(a, b)))$ time.
    \item<2-> Code:
        \begin{minted}{cpp}
          int lcm(int a, int b){
            return (a / gcd(a, b)) * b;
          }
        \end{minted}
    \item<3-> if $a = p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$ and $b = p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$, then:
        \[
          LCM(a, b) = p_1^{\max(e_1, f_1)} \cdot p_2^{\max(e_2, f_2)} \cdots p_k^{\max(e_k, f_k)}
        \]
  \end{itemize}
\end{frame}
\begin{frame}{1968A - Maximize?}
  \framesubtitle{Problem link: \url{https://codeforces.com/problemset/problem/1968/A}}
  \begin{itemize}
    \item<1-> You are given an integer \( x \). Your task is to find any integer \( y \) such that \( 1 \leq y < x \) and \( \gcd(x, y) + y \) is maximum possible. Note that if there is more than one \( y \) which satisfies the statement, you are allowed to find any.    
    \item<2-> $1 \leq t \leq 10^6$ and $1 \leq x \leq 10^{18}$
  \end{itemize}
\end{frame}
\begin{frame}{Rough}
  \framesubtitle{Rough - 1968A}
\end{frame}
\begin{frame}{Solution}
  \framesubtitle{Solution - 1968A}
  \begin{itemize}
    \item<1-> Answer is $x-1$.
  \end{itemize}
\end{frame}

\section{Modular Arithmetic}
\begin{frame}[fragile]{Modular Arithmetic}
  \begin{enumerate}
    \item $(a + b) \pmod m \equiv (a \mod m + b \mod m) \mod m$ 
    \item $(a - b) \pmod m \equiv (a \mod m - b \mod m) \mod m$
    \item $(a \cdot b) \pmod m \equiv (a \mod m \cdot b \mod m) \mod m$
    \item $a^b \pmod m \equiv (a \mod m)^b \mod m$
  \end{enumerate}
\end{frame}

\begin{frame}[fragile]{Modular Exponension}
  \begin{itemize}
    \item<1-> $a^b \pmod m$ can be computed in $\mathcal{O}(\log(b))$ time.
    \item<2-> Code:
        \begin{minted}{cpp}
          int mod_exp(int a, int b, int m){
            int res = 1;
            while (b > 0){
              if (b & 1)res = (res * a) % m;
              a = (a * a) % m;
              b >>= 1;
            }
            return res;
          }
        \end{minted}
  \end{itemize}
\end{frame}
\begin{frame}[fragile]{Mod Inverse}
  \begin{itemize}
    \item<1-> $a^{-1} \pmod m$ can be computed in $\mathcal{O}(\log(m))$ time.
    \item<2-> Code:
        \begin{minted}{cpp}
          int mod_inv(int a, int m){
            return mod_exp(a, m - 2, m);
          }
        \end{minted}
  \end{itemize}
\end{frame}
\begin{frame}[fragile]{NCR}
  \begin{itemize}
    \item<1-> $C(n, r) = \frac{n!}{r!(n - r)!}$
    \item<2-> $C(n, r) \pmod m$ can be computed in $\mathcal{O}(n)$ time.
    \item<3-> Code:
        \begin{minted}{cpp}
          int ncr(int n, int r, int m){
            if (r > n)return 0;
            return (fact[n] * mod_inv(fact[r], m) % m * mod_inv(fact[n - r], m) % m) % m;
          }
        \end{minted}
  \end{itemize}
\end{frame}

\input{mub.tex}

\begin{frame}{}
\framesubtitle{}
\begin{itemize}
    \item<1-> \begin{center}
        \large{Good Bye!}
    \end{center}
\end{itemize}

\end{frame}

\end{document}





